### 2013-02-15 doc/icml2013/algo.tex
edit

In this section, we describe our Spectral Experts algorithm
for estimating model parameters $\theta = (\pi, B)$.
The algorithm consists of two steps:
(i) low-rank regression to estimate certain symmetric tensors;
and (ii) tensor factorzation to recover the parameters.
The two steps can be performed efficiently using
convex optimization and tensor power method, respectively.
% Motivate how we can recover $B$ from regression by lower order moments
%ideally from lower order moments.
%Conventionally, when using the method of moments, the
%moments of observed variables are expressed algebraically in terms of
%the model parameters. Using these equations, we attempt to solve for the
%model parameters using empirical estimates of the moments.

@@

% Spectral
In recent years, there has a been substantial interest in ``spectral'' methods
for learning latent-variable models.  One line of work has
focused on observerable operator models \cite{hsu09spectral}
in which the true parameters are not recovered, but a reparametrization is used
which enables prediction.
Another line of work is based on the method of moments and uses eigendecomposition of a certain tensor
to recover the parameters \cite{anandkumar12moments,anandkumar12svd,AnandkumarHsuKakade2012}.
Our work extends this line of work into the space of discriminative models,
which requires regression to obtain the desired tensor.

% Unmixing
In spirit, Spectral Experts also bears some resemblence to the unmixing
algorithm for estimation of restricted PCFGs
\cite{hsu12identifiability}.
In that work, the observations (moments) provided a linear combination over
the compound parameters.  ``Unmixing'' involves solving for the compound
parameters by inverting a mixing matrix.  In this work,
each data point (appropriately transformed) provides a noisy slice of the
compound parameters; retrieving the compound parameters is more involved.

% Signal
The idea of performing low-rank regression on $y^2$ has been explored
in the context of signal recovery from magnitude measurements
\cite{candes11phaselift,ohlsson12phase}.
There, the actual observed response was $y^2$,
whereas in our case, we delibrately construct powers $y,y^2,y^3$
to identify the parameters.

@@

\begin{algorithm}[t]
  \caption{Spectral Experts}
  \label{algo:spectral-experts}
  \begin{algorithmic}[1]
    \INPUT Three independent datasets $\mathcal{D}_r = \{ (\xn{1}, \yn{1}), \cdots, (\xn{n}, \yn{n}) \}$ for $r = 1, 2, 3$;
    regularization strengths $\lambda_n^{(2)} = \frac{c_2}{\sqrt{n}}$, $\lambda_n^{(3)} = \frac{c_3}{\sqrt{n}}$;
    observation noise variance $\sigma^2$.
    \OUTPUT Parameters $\hat\theta = (\hat \pi, [\hat \beta_1 \mid \cdots \mid \hat \beta_k])$.
    \STATE Estimate compound parameters $M_2, M_3$ using \textbf{low-rank regression}:
    \begin{align}
      &\hat M_1 = \arg\min_{M_1} \label{eqn:estimateM1} \\
      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_1} (\innerp{M_1}{x} - y)^2, \nonumber \\
      &\hat M_2 = \arg\min_{M_2} \quad \lambda_n^{(2)} \|M_2\|_* + \label{eqn:estimateM2} \\
      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_2} (\innerp{M_2}{x\tp{2}} + \sigma^2 - y^2)^2, \nonumber \\
      &\hat M_3 = \arg\min_{M_3} \quad \lambda_n^{(3)} \|M_3\|_* + \label{eqn:estimateM3} \\
      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_3} (\innerp{M_3}{x\tp{3}} + 3 \sigma^2\innerp{\hat M_1}{x} - y^3)^2. \nonumber
    \end{align}
    \STATE Estimate the parameters $\theta = (\pi, B)$ using \textbf{tensor factorzation}:
    \begin{enumerate}
      \item [(a)] Compute whitening matrix $\hat W \in \Re^{d \times k}$ (such that $\hat W^\top
      \hat M_2 \hat W = I$) using SVD.
      \item [(b)] Compute the eigenvalues $\{\hat a_h\}_{h=1}^k$
      and eigenvectors $\{\hat v_h\}_{h=1}^k$
      of the whitened tensor $\hat M_3(\hat W, \hat W, \hat W) \in \Re^{k \times k \times k}$
      by using the robust tensor power method.
    \item [(c)] Set parameter estimates $\hat\pi_h = \frac{1}{\hat a_h^2}$
    and $\hat\beta_h = (\hat W^{\top})^\dagger (\hat a_h \hat v_h)$.
    \end{enumerate}
  \end{algorithmic}
\end{algorithm}

>>>>

In this section, we describe our Spectral Experts algorithm
for estimating model parameters $\theta = (\pi, B)$.
The algorithm consists of two steps:
(i) low-rank regression to estimate certain symmetric tensors;
and (ii) tensor factorization to recover the parameters.
The two steps can be performed efficiently using
convex optimization and tensor power method, respectively.
% Motivate how we can recover $B$ from regression by lower order moments
%ideally from lower order moments.
%Conventionally, when using the method of moments, the
%moments of observed variables are expressed algebraically in terms of
%the model parameters. Using these equations, we attempt to solve for the
%model parameters using empirical estimates of the moments.

@@

% Spectral
In recent years, there has a been substantial interest in ``spectral'' methods
for learning latent-variable models.  One line of work has
focused on observable operator models \cite{hsu09spectral}
in which the true parameters are not recovered, but a reparametrization is used
which enables prediction.
Another line of work is based on the method of moments and uses eigendecomposition of a certain tensor
to recover the parameters \cite{anandkumar12moments,anandkumar12svd,AnandkumarHsuKakade2012}.
Our work extends this line of work into the space of discriminative models,
which requires regression to obtain the desired tensor.

% Unmixing
In spirit, Spectral Experts also bears some resemblance to the unmixing
algorithm for estimation of restricted PCFGs
\cite{hsu12identifiability}.
In that work, the observations (moments) provided a linear combination over
the compound parameters.  ``Unmixing'' involves solving for the compound
parameters by inverting a mixing matrix.  In this work,
each data point (appropriately transformed) provides a noisy slice of the
compound parameters; retrieving the compound parameters is more involved.

% Signal
The idea of performing low-rank regression on $y^2$ has been explored
in the context of signal recovery from magnitude measurements
\cite{candes11phaselift,ohlsson12phase}.
There, the actual observed response was $y^2$,
whereas in our case, we deliberately construct powers $y,y^2,y^3$
to identify the parameters.

@@

\begin{algorithm}[t]
  \caption{Spectral Experts}
  \label{algo:spectral-experts}
  \begin{algorithmic}[1]
    \INPUT Three independent datasets $\mathcal{D}_r = \{ (\xn{1}, \yn{1}), \cdots, (\xn{n}, \yn{n}) \}$ for $r = 1, 2, 3$;
    regularization strengths $\lambda_n^{(2)} = \frac{c_2}{\sqrt{n}}$, $\lambda_n^{(3)} = \frac{c_3}{\sqrt{n}}$;
    observation noise variance $\sigma^2$.
    \OUTPUT Parameters $\hat\theta = (\hat \pi, [\hat \beta_1 \mid \cdots \mid \hat \beta_k])$.
    \STATE Estimate compound parameters $M_2, M_3$ using \textbf{low-rank regression}:
    \begin{align}
      &\hat M_1 = \arg\min_{M_1} \label{eqn:estimateM1} \\
      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_1} (\innerp{M_1}{x} - y)^2, \nonumber \\
      &\hat M_2 = \arg\min_{M_2} \quad \lambda_n^{(2)} \|M_2\|_* + \label{eqn:estimateM2} \\
      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_2} (\innerp{M_2}{x\tp{2}} + \sigma^2 - y^2)^2, \nonumber \\
      &\hat M_3 = \arg\min_{M_3} \quad \lambda_n^{(3)} \|M_3\|_* + \label{eqn:estimateM3} \\
      &\quad\quad\frac{1}{2n} \sum_{(x,y) \in \sD_3} (\innerp{M_3}{x\tp{3}} + 3 \sigma^2\innerp{\hat M_1}{x} - y^3)^2. \nonumber
    \end{align}
    \STATE Estimate the parameters $\theta = (\pi, B)$ using \textbf{tensor factorization}:
    \begin{enumerate}
      \item [(a)] Compute whitening matrix $\hat W \in \Re^{d \times k}$ (such that $\hat W^\top
      \hat M_2 \hat W = I$) using SVD.
      \item [(b)] Compute the eigenvalues $\{\hat a_h\}_{h=1}^k$
      and eigenvectors $\{\hat v_h\}_{h=1}^k$
      of the whitened tensor $\hat M_3(\hat W, \hat W, \hat W) \in \Re^{k \times k \times k}$
      by using the robust tensor power method.
    \item [(c)] Set parameter estimates $\hat\pi_h = \frac{1}{\hat a_h^2}$
    and $\hat\beta_h = (\hat W^{\top})^\dagger (\hat a_h \hat v_h)$.
    \end{enumerate}
  \end{algorithmic}
\end{algorithm}

### 2013-02-15 doc/icml2013/intro.tex
edit

% Discriminative models work better in practice Discriminative models
% with latent variables add expressive power (examples)
Discriminative latent-variable models,
which combine the high accuracy of discriminative models
with the compact expressiveness of latent-variable models,
have been widely applied to many tasks, including
object recognition \cite{quattoni04crf},
human action recognition \cite{wang09crf},
syntactic parsing \cite{petrov08discriminative},
and machine translation \cite{liang06discrimative}.
However, parameter estimation in these models is difficult;
past approaches have relied exclusively on local optimization (EM, 
gradient descent) and are vulerable to local optima.

@@

Let $[n] = \{ 1, \dots, n \}$ denote the first $n$ positive integers.

@@

% Norms
For a matrix $X$,
let $\|X\|_*$ denote the nuclear (trace) norm (sum of singular values),
let $\|X\|_F$ denote the Frobenius norm (square root of sum of squares of singular values),
let $\|X\|_\text{op}$ denote the operator norm (largest singular value).
For a rank-$K$ tensor $X$,
let $\|X\|_* = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_*$ denote
the average nuclear norm over all $p$ unfoldings,
and let $\|X\|_\text{op} = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_\text{op}$
denote the average operator norm over all $p$ unfoldings.

>>>>

% Discriminative models work better in practice Discriminative models
% with latent variables add expressive power (examples)
Discriminative latent-variable models,
which combine the high accuracy of discriminative models
with the compact expressiveness of latent-variable models,
have been widely applied to many tasks, including
object recognition \cite{quattoni04crf},
human action recognition \cite{wang09crf},
syntactic parsing \cite{petrov08discriminative},
and machine translation \cite{liang06discrimative}.
However, parameter estimation in these models is difficult;
past approaches have relied exclusively on local optimization (EM, 
gradient descent) and are vulnerable to local optima.

@@

Let $[n] = \{ 1, \dots, n \}$ denote the first $n$ positive integers.
We use $O(f(n))$ to denote a function $g(n)$ such that $\lim_{n \to\infty} g(n)/f(n) < \infty$.
We will only hide universal constants, exposing all dependence on the number of samples,
dimension, norms, noise variance, etc.

@@

% Norms
For a matrix $X$,
let $\|X\|_*$ denote the nuclear (trace) norm (sum of singular values),
let $\|X\|_F$ denote the Frobenius norm (square root of sum of squares of singular values),
let $\|X\|_\text{op}$ denote the operator norm (largest singular value).
For a rank-$K$ tensor $X$,
let $\|X\|_* = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_*$ denote
the average nuclear norm over all $p$ unfoldings,
and let $\|X\|_\text{op} = \frac{1}{p} \sum_{i=1}^p \|X_{(i)}\|_\text{op}$
denote the average operator norm over all $p$ unfoldings.
Let $\sigma_\text{min}(X)$ be the smallest singular value of $X$.

### 2013-02-15 doc/icml2013/model.tex
Minor fixes
edit

The mixture of linear regressions model \citep{VieleTong2002} defines
a conditional distribution over a response $y \in \Re$
given covariates $x \in \Re^d$.
The generation of $y$ given $x$ involves two steps:
(i) draw the mixture component $h \in [K]$ according to mixture proportions
$\pi = (\pi_1, \dots, \pi_K)$;
and (ii) generate $y$ from a Gaussian centered at $\beta_h^\top x$ with
variance $\sigma^2$.
\begin{eqnarray*}
  h &\sim& \mult(\pi), \\
  y \mid h &\sim& \normal{ \beta_{h}^T x}{ \sigma^2 }.
\end{eqnarray*}
The parameters of the model are $\theta = (\pi, B)$,
where $\pi \in \Re^d$ are the mixture proportions and
$B = [\beta_1 \mid \dots \mid \beta_K] \in \Re^{d \times K}$
are the regression coefficients.
We assume the variance of the observation noise $\sigma^2$ is known.
The learning problem is stated as follows:
given $n$ i.i.d. samples $(\xn{1}, \yn{1}), \dots, (\xn{n}, \yn{n})$
drawn from the model with some unknown parameters $\theta^*$,
return an estimate of the parameters $\hat\theta$.

% History of model
The mixture of linear regressions model has been applied
in the statistics literature for modeling music perception, where $x$ is the
actual tone and $y$ is the tone perceived by a musician \cite{VieleTong2002}.
% Reference mixture of experts.
% Acknowledge that mixture proportions % can't depend on $x$
The model is an instance of hierarchical mixture of experts
\cite{jacobs91experts}, in which the mixture proportions are allowed to depend
on $x$, known as a gating function.
This dependence allow the experts to be localized in input space,
providing more flexbility, but we do not consider this dependence in our model.

>>>>

The mixture of linear regressions model \citep{VieleTong2002} defines
a conditional distribution over a response $y \in \Re$
given covariates $x \in \Re^d$.
The generation of $y$ given $x$ involves two steps:
(i) draw the mixture component $h \in [K]$ according to mixture proportions
$\pi = (\pi_1, \dots, \pi_K)$;
and (ii) generate $y$ from a Gaussian centered at $\beta_h^\top x$ with
variance $\sigma^2$.
\begin{eqnarray*}
  h &\sim& \mult(\pi), \\
  y \mid h &\sim& \normal{ \beta_{h}^T x}{ \sigma^2 }.
\end{eqnarray*}
The parameters of the model are $\theta = (\pi, B)$,
where $\pi \in \Re^d$ are the mixture proportions and
$B = [\beta_1 \mid \dots \mid \beta_K] \in \Re^{d \times K}$
are the regression coefficients.
We assume the variance of the observation noise $\sigma^2$ is known and that the mixture components are independent.
The learning problem is stated as follows:
given $n$ i.i.d. samples $(\xn{1}, \yn{1}), \dots, (\xn{n}, \yn{n})$
drawn from the model with some unknown parameters $\theta^*$,
return an estimate of the parameters $\hat\theta$.

% History of model
The mixture of linear regressions model has been applied
in the statistics literature for modeling music perception, where $x$ is the
actual tone and $y$ is the tone perceived by a musician \cite{VieleTong2002}.
% Reference mixture of experts.
% Acknowledge that mixture proportions % can't depend on $x$
The model is an instance of hierarchical mixture of experts
\cite{jacobs91experts}, in which the mixture proportions are allowed to depend
on $x$, known as a gating function.
This dependence allow the experts to be localized in input space,
providing more flexibility, but we do not consider this dependence in our model.

### 2013-02-15 doc/icml2013/regression.tex
edit

Finally, $\eta_3(x)$ contains five terms plus an additional bias since the estimator uses $\hat M_1$ rather than $M_1$.
There are two additional considerations.
First, to bound $\hat\E_3[\epsilon^3]$, we use Lemma 19 of \cite{hsu13spherical}.
Second, the additional bias is small by the previous bound (and it is constructed on $\sD_1$, which is independent of $\sD_3$
used to contruct $\hat M_3$.
Therefore, we incur another $O(\sigma^2 \cdot R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
Handling the other terms using conventional means
$\|\hat\E_3[\eta_3(x) x\tp{3}]\|_\op = O(R^3 (L^3 R^3 + \sigma L^2 R^2 + \sigma^2 L R + \sigma^3 + \sigma^3 R) \sqrt{\frac{\log^3(1/\delta_1)}{n}})$.
Finally, taking $\delta_1 = \delta/3$, and taking the union bound,
we get our result, where we are bounding quantities fairly crudely for the sake of simplicity.
\end{proof}

>>>>

Finally, $\eta_3(x)$ contains five terms plus an additional bias since the estimator uses $\hat M_1$ rather than $M_1$.
There are two additional considerations.
First, to bound $\hat\E_3[\epsilon^3]$, we use Lemma 19 of \cite{hsu13spherical}.
Second, the additional bias is small by the previous bound (and it is constructed on $\sD_1$, which is independent of $\sD_3$
used to construct $\hat M_3$.
Therefore, we incur another $O(\sigma^2 \cdot R \sigma \sqrt{\frac{\log(1/\delta_1)}{n}})$.
Handling the other terms using conventional means
$\|\hat\E_3[\eta_3(x) x\tp{3}]\|_\op = O(R^3 (L^3 R^3 + \sigma L^2 R^2 + \sigma^2 L R + \sigma^3 + \sigma^3 R) \sqrt{\frac{\log^3(1/\delta_1)}{n}})$.
Finally, taking $\delta_1 = \delta/3$, and taking the union bound,
we get our result, where we are bounding quantities fairly crudely for the sake of simplicity.
\end{proof}
